Measuring Perceptions and Preferences for Meritocracy

### 2020-09-10 docs/paper.Rmd
correcciones finales en texto

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
abstract: "Economic and social inequality have raised growing concerns and crises across societies. One of social science concepts associated to the maintenance of inequality is the belief in meritocracy, which would legitimize economic differences based on effort and talent. Despite its wide use,  empirical research on meritocracy is something relatively novel. A number of studies have relied mostly in secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items. Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and measurement of meritocracy. Based on this critical analysis, we propose  an items battery called *Perceptions and preferences for meritocracy scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141). The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions. The discussion highlights the relevance of considering these different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::pdf_document2:
    template: null
    toc: false 
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography:
  - ../input/bib/EconomiaMoral.bib
  - ../input/bib/meritocracy.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

@@

Based on the previous limitations and confussions in the measurement of meritocracy presented in the previous section, in this paper we propose and test an instrument with the following characteristics:

@@

The data was obtained through an online questionnaire which was part of a larger study on meritocracy and preferences developed in Chile in 2019 and funded by the national scientific agency FONDECYT. The questionnaire was programmed in Qualtrics and the fieldwork was in charge of an external online survey agency ([netquest.cl](www.netquest.cl)) during December 2019 and January 2020. The sample was selected from a non-probabilistic design in three large cities in Chile. The quota method based on age, sex and educational level was used. Quotas were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country and is also the counterpart agency of ISSP (International Social Survey Programme) in Chile. A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale and those who did not accept informed consent. As it usually occurs online samples, there were some limitations in achieving the quotas for lower educational levels.

@@

CFA was conducted using the `lavaan` package (version 0.6-3; Rosseel, 2020) with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020). As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio:> 3).

A pre-registration was made in the OSF platform, [Perceptions and Preferences for Meritocracy Scale
Public registration](https://osf.io/z45y2) In it, it was indicated, before carrying out the analyzes, the four-dimensional theoretical model, the level of measurement of the variables, the statistical tests to be used with their respective evaluation parameters, along with other important aspects of the research design.

@@

Como puede observarse en la tabla \@ref(tab:desc01), los indicadores poseen valores que van desde el 1 (totalmente en desacuerdo) al 5 (totalmente de acuerdo). Se observan promedios desde 2.41 correspondiente a preferencia-contactos, hasta 3.89 correspondiente a la preferencia-esfuerzo. Ambos indicadores son coherentes con la adhesión general a la meritocracia, privilegiando aspectos individuales como el esfuerzo.

@@

A partir del la Figura \@ref(fig:plotlikert) se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto. En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento.

@@

En la Figura \@ref(fig:corpoly) se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia esfuerzo con talento r=0.56), mientras que se observan correlaciones de baja intensidad entre el resto de los cruces. Destacan además las relaciones entre las percepciones y preferencias de tipo meritocráticas, lo que no ocurre con los indicadores no meritocraticos.

@@

En suma, los análisis descriptivos señalan una relativa adhesión a la moral meritocratica, las cuales, a modo general, se expresan en una mayor preferencia por criterios meritocraticos y una menor por no meritocráticos. Igualmente, es observable una relativamente baja percepción de meritocracia. Se observa además una relación coherente entre los indicadores según lo propuesto en el modelo teórico en el preregistro del estudio, es decir, los pares de items asociados a un factor específico muestran correlaciones con un tamaño de efecto grande (por ejemplo, preferencias meritocrácticas por los items asociados a esfuerzo y a talento). En particular las asociaciones entre esfuerzo y talento son relevantes, ya que desestiman previos supuestos sobre que el talento no sería un criterio meritocrático [@mijs_paradox_2019], de otra manera la correlación sería cero o negativa. Junto a ello, vemos que no existe una correlacion negativa entre aspectos meritocráticos y no meritocráticos, desestimando los supuestos de estudios previos que  señalaban que estas dimensiones serían los polos opuestos de un mismo continuo [@reynolds_perceptions_2014].

@@

This section tests the item's battery and the corresponding conceptual framework under test. For this, we first estimate a confirmatory factor analysis model for the whole sample, and secondly we test the order effects applying the same model to each of the three order versions.

@@

In \@ref(fig:meas02) the results of the estimation for the four-factor model with complete sample. This model have a good fit (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88. Furthermore, considering the correlation between the latent variables, it is observed that the most related variables are meritocratic perceptions and preferences, which have a positive correlation of 0.59, while non-meritocratic perceptions and preferences do not present a considerable correlation.

@@

Como se puede observar en la Tabla \@ref(tab:tab-loads), todos los modelos, independiente del grupo y el órden de los items, obtuvieron un ajuste adecuado. con CFI superiores a .95 y RMSEA inferiores a .8, por contraparte ni un modelo logró un chi-square no significativo, aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3. El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, $\chi2$(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001).  Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968, RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).

@@

EL modelo teórico propuesto de cuatro factores ajustó, como se observa en la tabla 4, mejor que el modelo de contraste de 1 factor. El modelo teórico ajustó de manera relativamente adecuada, pues muestra indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque posee indicadores deficientes para la prueba $\chi2$(df=14)=67.6, p-value=.000. Para evaluar posibles mejoras de la escala, se analizaron las relaciones propuestas por los índices de modificación. Estos indican la existencia de dos cargas cruzadas no especificadas. Cuando se generó un modelo siguiendo estas recomendaciones, hubo una mejora considerable del modelo, aunque el nuevo modelo tampoco obtuvo un $\chi2$ ratio menor a 3 y obtuvo cargas factoriales muy bajas ($\tau$)< 0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo aceptar las propuesta de los índices de modificación cuando se posee teoria y evidencia sólida, se ha decidido no incorporar estos parámetros al modelo.

>>>>

---
title: "Measuring Perceptions and Preferences for Meritocracy"
css: "custom.css"
linestretch: '1.5'
link-citations: yes
author: "Juan Carlos Castillo, Julio Iturra & Francisco Meneses <br><br>"
abstract: "Economic and social inequality have raised growing concerns and crises across societies. One of social science concepts associated to the maintenance of inequality is the belief in meritocracy, which would legitimize economic differences based on effort and talent. Despite its wide use, empirical research on meritocracy is something relatively novel. A number of studies have relied mostly in secondary data to operationalize meritocracy, with a large variation in the use and interpretation of survey items. Starting from a review of studies that measure meritocracy, this article identifies a series of drawbacks and inconsistencies within and between studies regarding the conceptualization and indicators of meritocracy. Based on this critical analysis, we propose an item's battery called *Perceptions and preferences for meritocracy scale*, which is tested with confirmatory factor analysis with data from an online survey study (N=2,141). The results support the proposed conceptual structure which not only distinguishes between perceptions and preferences, but also between meritocratic and non-meritocratic dimensions. The discussion highlights the relevance of considering these different dimensions in order to advance in the study of meritocracy."
output:
  bookdown::pdf_document2:
    template: null
    toc: false
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography:
  - ../input/bib/EconomiaMoral.bib
  - ../input/bib/meritocracy.bib
editor_options:
  chunk_output_type: console
geometry: margin=0.78in
header-includes:
  # - \usepackage[spanish,es-tabla,es-nodecimaldot]{babel}
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
---

@@

Based on the previous limitations and confusions in the measurement of meritocracy presented in the previous section, in this paper we propose and test an instrument with the following characteristics:

@@

The data was obtained through an online questionnaire which was part of a larger study on meritocracy and preferences developed in Chile in 2019 and funded by the national scientific agency FONDECYT. The questionnaire was programmed in Qualtrics and the fieldwork was in charge of an external online survey agency ([netquest.cl](www.netquest.cl)) during December 2019 and January 2020. The sample was selected from a non-probabilistic quota design in three large cities in Chile. The quotas were generated based on the survey of the Public Studies Center (CEP, 2019), which has a high prestige in the country and is also the counterpart agency of ISSP (International Social Survey Programme) in Chile. A total sample of 2,141 people was collected, excluding those who did not answer the questions on the scale and those who did not accept informed consent. As it usually occurs online samples, there were some limitations in achieving the quotas for lower educational levels.

@@

CFA was conducted using the `lavaan` R package (version 0.6-3; Rosseel, 2020) with diagonally weighted least squares (DWLS) estimation due to the items’ ordinal level of measurement (Kline, 2016; Rosseel, 2020). As recommended by Brown (2008), we assessed model fit by jointly considering the comparative fit index and Tucker-Lewis Index (CFI and TLI; acceptable fit > 0.95), Root of the average squared residual approximation (RMSEA; acceptable fit < 0.08),  Chi-square: (p-value; acceptable fit > 0.05, and Chi-square ratio:> 3).

A pre-registration was made in the OSF platform, available in the following link: [https://osf.io/z45y2](https://osf.io/z45y2). In this pre-registration are the hypothesis regarding the four-dimensional conceptual model underlying the scale,  the variables measurement levels, the statistical tests to be performed with their respective evaluation parameters, along with other important aspects of the research design.

@@

Como puede observarse en la tabla \@ref(tab:desc01), los indicadores poseen valores que van desde el 1 (totalmente en desacuerdo) al 5 (totalmente de acuerdo). Se observan promedios desde 2.41 correspondiente a preferencia-contactos, hasta 3.89 correspondiente a la preferencia-esfuerzo. Ambos indicadores son coherentes con la adhesión general a la meritocracia reportada en estudios previos, privilegiando aspectos individuales como el esfuerzo.

@@

Los gráficos presentados en la Figura \@ref(fig:plotlikert) presentan información desagregada y comparable de las distintas categorías de respuesta de cada item.  Se observa que en general hay una mayor percepción de elementos no meritocráticos que meritocráticos, mientras que en el caso de preferencias ocurre lo opuesto. En cuanto a las preferencias, llama la atención el rol preponderante del esfuerzo por sobre el talento como criterio de preferencia meritocrático.

@@

En la Figura \@ref(fig:corpoly) se observan relaciones de moderada a alta intensidad entre los indicadores que corresponden al mismo factor (por ejemplo, percepción de meritocracia por esfuerzo y por talento, r=0.56). Destacan además las relaciones entre las percepciones y preferencias de tipo meritocráticas, lo que no ocurre con los indicadores no meritocraticos.

@@

En suma, los análisis descriptivos señalan una relativa adhesión a la moral meritocratica, las cuales se expresan en una mayor preferencia por criterios meritocraticos y una menor por no meritocráticos, aún cuando la preferencia por elementos no-meritocraticos alcanza una proporción de más de 20%. Igualmente, es observable una relativamente baja percepción de meritocracia. Se observa además una relación coherente entre los indicadores según lo propuesto en el modelo teórico en el preregistro del estudio, es decir, los pares de items asociados a un factor específico muestran correlaciones con un tamaño de efecto grande (por ejemplo, preferencias meritocrácticas por los items asociados a esfuerzo y a talento). En particular las asociaciones entre esfuerzo y talento son relevantes, ya que desestiman previos supuestos sobre que el talento no sería un criterio meritocrático [@mijs_paradox_2019], de otra manera la correlación sería cero o negativa. Junto a ello, vemos que no existe una correlación negativa entre aspectos meritocráticos y no meritocráticos, desestimando los supuestos de estudios previos que  señalaban que estas dimensiones serían los polos opuestos de un mismo continuo [@reynolds_perceptions_2014].

@@

This section estimates the fit of the conceptual model behind the Perceptions and Preferences for Meritocracy Scale. For this, we first estimate a confirmatory factor analysis model for the whole sample, and secondly we test the order effects applying the same model to each of the three order versions.

@@

\@ref(fig:meas02) SHOWS the results of the estimation for the four-factor model with complete sample. This model presents adequate fit indicators (CFI = 0.977, RMSEA = 0.079, $\chi2$(df = 14) = 28,03, p = 0.014) and it presents factorial loads with weights that vary between 0.63 and 0.88. Furthermore, considering the correlation between the latent variables, it is observed that the variables with the highest correlations are meritocratic perceptions and preferences, which have a positive correlation of 0.59, while non-meritocratic perceptions and preferences do not present a considerable correlation.

@@

Como se puede observar en la Tabla \@ref(tab:tab-loads), todos los modelos, independiente del grupo y el órden de los items, obtuvieron un ajuste adecuado. con CFI superiores a .95 y RMSEA inferiores a .8. Sin embargo ningún modelo logró un chi-square no significativo (algo esperable en muestras grandes), aunque tanto el modelo aleatorio como el primero obtuvieron un adecuado chi-square ratio menor a 3. El primero de los órdenes fue el que obtuvo mejor ajuste (CFI=0.998, TLI= 0.995,RMSEA=0.037, $\chi2$(df=14)=28,03, p = 0.014), seguido por el orden aleatorio de los ítems (CFI=0.992, TLI=0.984,RMSEA=0.051, $\chi2$(df=14)=39.09, p < 0.001). Por su parte, la escala ordenada por temáticas parece generar un efecto framing en el cual la relación entre las percepciones y las preferencias parecen sobreestimadas, afectando de esta manera el ajuste (CFI=0.984, TLI=0.968, RMSEA=.071, $\chi2$(df=14)=64.156, p < 0.001).

@@

EL modelo teórico propuesto de cuatro factores ajustó, como se observa en la tabla 4, mejor que el modelo de contraste de 1 factor. El modelo teórico ajustó de manera relativamente adecuada, pues muestra indicadores óptimos para CFI= 0.987, TLI = 0.975 y RMSEA=.041, aunque posee indicadores deficientes para la prueba $\chi2$(df=14)=67.6, p-value=.000. Para evaluar posibles mejoras de la escala, se analizaron las relaciones propuestas por los índices de modificación. Estos indican la existencia de dos cargas cruzadas no especificadas. Cuando se generó un modelo siguiendo estas recomendaciones, hubo una mejora considerable del modelo, aunque el nuevo modelo tampoco obtuvo un $\chi2$ ratio menor a 3 y obtuvo cargas factoriales muy bajas ($\tau$)< 0.15, por lo tanto, siguiendo las recomendación de Brown (2006) de solo aceptar las propuesta de los índices de modificación cuando se posee teoría y evidencia sólida, se ha decidido no incorporar estos parámetros al modelo.
